---
title: "Covid91 vaccine study Final2023"
author: "Chris Gonzalez"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <-FinalDat %>% mutate(infected = recode(infected, aids = "covid", none = "OK" )) %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
FinalData$treatment <- relevel(FinalData$treatment,ref="placebo")
#FinalData$infected <- relevel(FinalData$infected,ref="OK")
male_data = filter(FinalData, sex =="M")
female_data = filter(FinalData, sex =="F")
gay_data = filter(FinalData, LGBTQ =="gay")
drug_data = filter(FinalData, DrugUser =="yes")
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Methods

The project aims to evaluate the efficacy of the new preventative vaccine, Covid921, through clinical trials. Considering potential variations in effectiveness across different sub populations, which includes male,female, LGBTQ, and drug users, a comprehensive report will be prepared for each subgroup, including null and alternative hypothesis statements and proper statistical analysis of p-values. The methods employed include visual analysis using bar charts to gain initial insights, followed by numerical examination using Pearson's Chi-squared test with Yates for a table analysis. Lastly, inferential results will be obtained using Fisher's test to derive a more comprehensive understanding and draw conclusions within each subgroup. Based on these analysis, appropriate recommendations regarding vaccine uptake will be provided for each population segment.


## Males
### Graphical Descriptive Results
```{r}
dd2 <- male_data %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```




## Females


## LGBTQ



## Druggies


# Overall Results and Conclusions