---
title: "Covid91 vaccine study Final2023"
author: "Chris Gonzalez"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <-FinalDat %>% mutate(infected = recode(infected, aids = "covid", none = "OK" )) %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
FinalData$treatment <- relevel(FinalData$treatment,ref="placebo")
#FinalData$infected <- relevel(FinalData$infected,ref="OK")
male_data = filter(FinalData, sex =="M")
female_data = filter(FinalData, sex =="F")
gay_data = filter(FinalData, LGBTQ =="gay")
drug_data = filter(FinalData, DrugUser =="yes")
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Methods

The project aims to evaluate the efficacy of the new preventative vaccine, Covid921, through clinical trials. Considering potential variations in effectiveness across different sub populations, which includes male,female, LGBTQ, and drug users, a comprehensive report will be prepared for each subgroup, including null and alternative hypothesis statements and proper statistical analysis of p-values. The methods employed include visual analysis using bar charts to gain initial insights, followed by numerical examination using Pearson's Chi-squared test with Yates for a table analysis. Next, inferential results will be obtained using Fisher's test to derive a more comprehensive understanding and draw conclusions within each subgroup. Finally, I will employ an effectiveness confidence interval formula along with coding to calculate the vaccine's efficacy within each subgroup. Based on these analysis, appropriate recommendations regarding vaccine uptake will be provided for each population segment.


## Males
### Graphical Descriptive Results
```{r}
dd2 <- male_data %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

Upon visual examination of the data represented in a box plot, it becomes evident that male patients who received a placebo exhibited a higher rate of covid19 infections, accounting for approximately 69% of male infections. In contrast, male patients who received the vaccine had a lower infection rate, constituting approximately 30% of the male infections. This graphical representation highlights the disparity in infection rates between the two treatment groups and suggests a potential association between the treatment (placebo vs. vaccine) and covid19 infection susceptibility in male patients. However, further rigorous statistical analysis would be necessary to establish the significance of these findings and draw firm conclusions about the vaccine's effectiveness in mitigating covid19 infections in this specific male population.

### Numerical Descriptive Results

```{r}
table1 <- xtabs(~infected + treatment, data=male_data)
rowPerc(table1)
colPerc(table1)
```
The data presented in the contingency table shows the distribution of individuals based on treatment (placebo and vaccine) and infection status covid19 and OK (uninfected). Out of the total male individuals in the study, 70% were in the placebo group, and 30% were in the vaccine group contracted covid19. Among those in the placebo group the males studies were 2.55 times more chance of being infected with covid19. Overall, the table indicates a higher proportion of covid19 infections in the placebo group that were 2.55 times more likely to contract covid19, suggesting a potential association between the treatment and infection status. However, a thorough statistical analysis, such as chi-square tests or logistic regression, would be required to determine the significance of this association and draw more definitive conclusions about the effectiveness of the vaccine in preventing covid19 infections.

### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
malefish <- fisher.test(table1)
malefish
```

The analysis was conducted using two statistical tests to examine the association between treatment and infection status in a study of males. Firstly, Pearson's Chi-squared test with Yates' continuity correction was performed, yielding a significant chi-square statistic of an extremely small p-value = 0.0000000000001098. Secondly, Fisher's Exact Test for Count Data was employed, producing an equally significant p-value = 0.00000000000004631. The estimated odds ratio of 2.242522 suggests that individuals in the vaccine group had approximately 2.24 times higher odds of being infected compared to those in the placebo group. These findings collectively indicate a clear and substantial relationship between the treatment and infection status, highlighting the potential impact of the vaccine on susceptibility to infection.

### Effectiveness Confidence Interval 

```{r}
effect <- malefish$p.value/(malefish$p.value +1)
effectivenessmale = malefish$estimate/(malefish$estimate +1)*100
print(effectivenessmale)
```
```{r}
low=(malefish$conf.int[1]/(malefish$conf.int[1] +1))*100
hi=(malefish$conf.int[2]/(malefish$conf.int[2] +1))*100
options(digits=2)
```

#### Effectiveness Confidence Interval `r low`  to `r hi`
Based on the analysis of male data, the treatment demonstrated an effectiveness of 69.2% (95% CI: 64.26% - 73.76%). These results indicate a notable efficacy of the treatment among male participants due to the effectiveness being above 50%, with a substantial proportion experiencing positive outcomes. The narrow confidence interval suggests a high level of precision in our estimation, providing strong confidence that the true effectiveness of the treatment falls within the reported range. These findings offer valuable insights into the treatment's potential impact on the male demographic and contribute significantly to our understanding of its effectiveness in this specific population.


## Females
### Graphical Descriptive Results
```{r}
dd2 <- female_data %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

The box plot analysis, focusing on female patients receiving either the vaccine or placebo, reveals some intriguing findings. Among female patients who received the placebo, approximately 60% were infected with the covid19 virus, while those who received the vaccine had a lower infection rate of around 40%. Interestingly, both the placebo and vaccine groups showed a similar proportion of patients who remained uninfected, at approximately 50%. These observations raise questions about the factors influencing the infection rates in both groups and warrant further in-depth analysis to gain a clearer understanding of the underlying dynamics. Additional investigations and careful consideration of potential confounding variables are essential to elucidate the factors contributing to the differential infection rates between the placebo and vaccine groups in the female population.

### Numerical Descriptive Results
```{r}
table2 <- xtabs(~infected + treatment, data=female_data)
rowPerc(table2)
colPerc(table2)
```
The updated data presented in the contingency table pertains to female individuals and their infection status based on treatment (placebo and vaccine). Among female participants, 60% were infected with covid19 in the placebo group, while 40% were infected in the vaccine group. The overall distribution shows an equal split, with 50% of female individuals in each treatment group belonging to the OK category.

Analyzing the percentages within each treatment group, the females in the placebo group had 2 times chance of getting infected with covid19, leaving 98% uninfected. In contrast, in the vaccine group, it was shown to be 1.4 times less likely for females patients to be infected with covid19. These findings suggest a higher proportion of covid19 infections among females in the placebo group compared to those in the vaccine group.

However, to assess the significance of this discrepancy and establish any potential causal relationship between the treatment and infection status specifically in the female population, further rigorous statistical analysis would be required. The observed data provides valuable initial insights, but drawing definitive conclusions would necessitate applying appropriate statistical tests to evaluate the significance of the observed differences.

### Inferential Results
```{r}
chisq.test(table2)
chisqtestGC(table2)
femalefish <- fisher.test(table2)
femalefish
```
In the female patients, a comprehensive statistical analysis was performed to investigate the relationship between treatment and infection status. Pearson's Chi-squared test with Yates' continuity correction was conducted on the data, revealing a significant association between the treatment (placebo and vaccine) and infection status (covid19 and OK) in females. The test yielded a chi-square statistic of 13 with 1 degree of freedom and an impressively low p-value= 0.0003, indicating strong evidence against the null hypothesis. Additionally, Fisher's Exact Test for Count Data was applied, producing a similarly small p-value= 0.0002 and a 95% confidence interval for the odds ratio between 1.2 and 1.9.

### Effectiveness Confidence Interval 

```{r}
effect <- femalefish$p.value/(femalefish$p.value +1)
effectivenessfemale = femalefish$estimate/(femalefish$estimate +1)*100
print(effectivenessfemale)
```

```{r}
low=(femalefish$conf.int[1]/(femalefish$conf.int[1] +1))*100
hi=(femalefish$conf.int[2]/(femalefish$conf.int[2] +1))*100
options(digits=2)
```
#### Effectiveness Confidence Interval `r low`  to `r hi`
The analysis of female data revealed a treatment effectiveness of 60% (95% CI: 54.53% - 65.07%). These results highlight a significant efficacy of the treatment among female participants due to the results being above 50%, with a substantial proportion experiencing positive outcomes. The narrow confidence interval indicates a high level of precision in our estimation, instilling strong confidence that the true effectiveness of the treatment lies within the reported range. These findings provide valuable insights into the treatment's potential impact on the male demographic and significantly contribute to our understanding of its effectiveness in this specific population.

## LGBTQ

### Graphical Descriptive Results
```{r}
dd2 <- gay_data %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

Upon initial observation of the box plot graph, a notable pattern emerges in the infection rates within the LGBTQ community based on the treatment received (placebo and vaccine). Among gay individuals who received the vaccine, the covid19 infection rate is approximately 85%, significantly higher than the 15% infection rate observed in those who received the placebo. An interesting observation is that among the individuals who remained uninfected, approximately 50% received the placebo, while the other 50% received the vaccine. This observation raises intriguing questions about the vaccine's effectiveness in reducing infection rates within the LGBTQ community.This can be due to a smaller sample size in patients studied causing a lower frequency. Further investigation and careful analysis are required to comprehensively understand the underlying factors contributing to these trend. 

### Numerical Descriptive Results

```{r}
table3 <- xtabs(~infected + treatment, data=gay_data)
rowPerc(table3)
colPerc(table3)
```

In the LGBTQ data presented in the contingency table relates to individuals' infection status based on treatment (placebo and vaccine). Among gay individuals, 15% were infected with covid19 in the placebo group, whereas 85% were infected in the vaccine group. In the OK (uninfected) category, 50% of gay individuals belonged to the placebo group, and 50% to the vaccine group. In the vaccine group, its shows to be 3.5 times more likely to be infected with covid19. These findings suggest a notably higher proportion of covid19 infections among gay individuals in the vaccine group compared to the placebo group. 

### Inferential Results
```{r}
chisq.test(table3)
chisqtestGC(table3)
gayfish <- fisher.test(table3)
gayfish
```

The data presented above indicates the results of a statistical analysis conducted on LGBTQ gay patients using Pearson's Chi-squared test with Yates' continuity correction. The test aimed to explore the relationship between the treatment (placebo or vaccine) and the occurrence of Covid19 infection (infected or OK). The Chi-squared test yielded a significant result  p-value = 0.00000007, suggesting that there is a statistically significant association between the treatment and the occurrence of Covid19 infection among LGBTQ gay patients. 

### Effectiveness Confidence Interval
```{r}
effect <- gayfish$p.value/(gayfish$p.value +1)
effectivenessgay = gayfish$estimate/(gayfish$estimate +1)*100
print(effectivenessgay)
```
```{r}
low=(gayfish$conf.int[1]/(gayfish$conf.int[1] +1))*100
hi=(gayfish$conf.int[2]/(gayfish$conf.int[2] +1))*100
options(digits=2)
```
#### Effectiveness Confidence Interval `r low`  to `r hi`
In the LGBTQ gay patient population, the data analysis demonstrated a treatment effectiveness of 15% (95% CI: 7.603% - 26.56%). As the effectiveness falls below 50%, it indicates that the vaccine is ineffective in providing substantial protection within this specific group. 


## Druggies

### Graphical Descriptive Results
```{r}
dd2 <- drug_data %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

At first glance of the data pertaining to drug users, it seems that the vaccinated group has a higher rate of infection, with 85% of vaccinated individuals getting infected, compared to only 10% of those who received the placebo. Among the uninfected, approximately 50% received the placebo, while 48% were vaccinated. This discrepancy raises more questions since the percentage of infected individuals is much higher in the vaccinated group compared to the placebo group. Further analysis of the data is required to draw proper conclusions.

### Numerical Descriptive Results

```{r}
table4 <- xtabs(~infected + treatment, data=drug_data)
rowPerc(table4)
colPerc(table4)
```
In the of drug users patients, the data presented in the contingency table reveals the infection status based on the treatment received (placebo and vaccine). Among drug users, 12% were infected with covid19 in the placebo group, while 88% were infected in the vaccine group. In the category of "OK" (uninfected), 51% of drug users belonged to the placebo group, and 49% to the vaccine group. Analyzing the percentages within each treatment group, 0.4% of drug users in the placebo group were infected with covid19, while 99.6% were not infected. In the vaccine group, the vaccinated were 3.1 times more likely to be infected with covid19. These findings suggest a considerably higher proportion of covid19 infections among drug users who received the vaccine compared to those who received the placebo. However, further in-depth analysis is necessary to assess the significance of these results and determine any potential associations between the treatment and infection status specifically in the drug user population. Understanding the factors influencing infection rates is crucial for tailoring effective prevention and intervention strategies in this vulnerable group.

### Inferential Results
```{r}
chisq.test(table4)
chisqtestGC(table4)
drugfish <- fisher.test(table4)
drugfish
```

The provided data subject are patients that use drugs. The Chi-squared test was used to analyze the association between treatment (placebo and vaccine) and infection rates (Covid19 and OK) among drug users. According to the results, the Chi-square statistic yielded a value of 35 with 1 degree of freedom and an extremely low p-value= 0.000000003. This indicates a highly significant association between the treatment and infection rates among drug users. The Fisher's Exact Test, which tests the odds ratio, resulted in a p-value= 0.0000000002. Overall, this statistical analysis provides strong evidence that there is a significant association between the treatment (placebo and vaccine) and infection rates (Covid19 and OK) among drug users.

### Effectiveness Confidence Interval
```{r}
effect <- drugfish$p.value/(drugfish$p.value +1)
effectivenessdrug = drugfish$estimate/(drugfish$estimate +1)*100
print(effectivenessdrug)
```

```{r}
low=(drugfish$conf.int[1]/(drugfish$conf.int[1] +1))*100
hi=(drugfish$conf.int[2]/(drugfish$conf.int[2] +1))*100
options(digits=2)
```
#### Effectiveness Confidence Interval `r low`  to `r hi`
In the drug user patients population, the data analysis demonstrated a treatment effectiveness of 11% (95% CI: 4.59% - 21.82%). As the effectiveness falls below 50%, it indicates that the vaccine is ineffective in providing substantial protection within this specific group. 



# Overall Results and Conclusions

In conclusion, the efficacy of the Covid19 treatment was found to be inadequate in safeguarding the LGBTQ and drug users' populations. The effectiveness of the treatment for these groups was reported as `r effectivenessgay` for gay individuals and `r effectivenessdrug` for drug users.

Conversely, the vaccine demonstrated notable success in protecting both women and men against Covid19. The effectiveness of the vaccine in males was `r effectivenessmale`, while in females, it was `r effectivenessfemale`. These gender-specific categories exhibited the highest level of protection with this vaccine.

However, concerning individuals who identified as gay, the vaccine displayed limited efficacy. Those who identified as gay and received the vaccine were found to be 3.5 times more likely to contract Covid19 compared to those who received the placebo.Similarly, vaccinated drug users also exhibited reduced effectiveness, being 3.1 times more likely to acquire Covid19 after receiving the vaccine.

As a result, if you belong to the LGBTQ community or are a drug user, it is not recommended to receive the vaccine due to its limited effectiveness. However, for individuals identifying as straight female or male, the vaccine has shown to be effective and a recommendation would be appropriate to protect against Covid19.





