---
title: "Covid91 vaccine study Final2023"
author: "Chris Gonzalez"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <-FinalDat %>% mutate(infected = recode(infected, aids = "covid", none = "OK" )) %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
FinalData$treatment <- relevel(FinalData$treatment,ref="placebo")
#FinalData$infected <- relevel(FinalData$infected,ref="OK")
male_data = filter(FinalData, sex =="M")
female_data = filter(FinalData, sex =="F")
gay_data = filter(FinalData, LGBTQ =="gay")
drug_data = filter(FinalData, DrugUser =="yes")
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Methods

The project aims to evaluate the efficacy of the new preventative vaccine, Covid921, through clinical trials. Considering potential variations in effectiveness across different sub populations, which includes male,female, LGBTQ, and drug users, a comprehensive report will be prepared for each subgroup, including null and alternative hypothesis statements and proper statistical analysis of p-values. The methods employed include visual analysis using bar charts to gain initial insights, followed by numerical examination using Pearson's Chi-squared test with Yates for a table analysis. Lastly, inferential results will be obtained using Fisher's test to derive a more comprehensive understanding and draw conclusions within each subgroup. Based on these analysis, appropriate recommendations regarding vaccine uptake will be provided for each population segment.


## Males
### Graphical Descriptive Results
```{r}
dd2 <- male_data %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

Upon visual examination of the data represented in a box plot, it becomes evident that male patients who received a placebo exhibited a higher rate of covid91 infections, accounting for approximately 69% of male infections. In contrast, male patients who received the vaccine had a lower infection rate, constituting approximately 30% of the male infections. This graphical representation highlights the disparity in infection rates between the two treatment groups and suggests a potential association between the treatment (placebo vs. vaccine) and covid91 infection susceptibility in male patients. However, further rigorous statistical analysis would be necessary to establish the significance of these findings and draw firm conclusions about the vaccine's effectiveness in mitigating covid infections in this specific male population.

### Numerical Descriptive Results

```{r}
table1 <- xtabs(~infected + treatment, data=male_data)
rowPerc(table1)
colPerc(table1)
```
The data presented in the contingency table shows the distribution of individuals based on treatment (placebo and vaccine) and infection status covid19 and OK (uninfected). Out of the total male individuals in the study, 69.57% were in the placebo group, and 30.43% were in the vaccine group. Among those in the placebo group, 2.55% were infected with covid19, and 97.45% were not infected (OK). In contrast, in the vaccine group, 1.15% were infected with covid19, and 98.85% were not infected. Overall, the table indicates a higher proportion of covid19 infections in the placebo group (2.55%) compared to the vaccine group (1.15%), suggesting a potential association between the treatment and infection status. However, a thorough statistical analysis, such as chi-square tests or logistic regression, would be required to determine the significance of this association and draw more definitive conclusions about the effectiveness of the vaccine in preventing covid19 infections.

## Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
malefish <- fisher.test(table1)
malefish
```

The analysis was conducted using two statistical tests to examine the association between treatment and infection status in a study of males. Firstly, Pearson's Chi-squared test with Yates' continuity correction was performed, yielding a significant chi-square statistic of 55.1832 with 1 degree of freedom and an extremely small p-value of 0.0000000000001098. Secondly, Fisher's Exact Test for Count Data was employed, producing an equally significant p-value of 0.00000000000004631 and a 95% confidence interval for the odds ratio of [1.798068, 2.810778]. The estimated odds ratio of 2.242522 suggests that individuals in the vaccine group had approximately 2.24 times higher odds of being infected compared to those in the placebo group. These findings collectively indicate a clear and substantial relationship between the treatment and infection status, highlighting the potential impact of the vaccine on susceptibility to infection.

### Effectiveness Confidence Interval 

```{r}
effect <- malefish$p.value/(malefish$p.value +1)
effectivenessmale = malefish$estimate/(malefish$estimate +1)*100
print(effectivenessmale)
```
```{r}
low=(malefish$conf.int[1]/(malefish$conf.int[1] +1))*100
hi=(malefish$conf.int[2]/(malefish$conf.int[2] +1))*100
options(digits=2)
```

#### Effectiveness Confidence Interval `r low`  to `r hi`
Based on the analysis of male data, the treatment demonstrated an effectiveness of 69.2% (95% CI: 64.26% - 73.76%). These results indicate a notable efficacy of the treatment among male participants, with a substantial proportion experiencing positive outcomes. The narrow confidence interval suggests a high level of precision in our estimation, providing strong confidence that the true effectiveness of the treatment falls within the reported range. These findings offer valuable insights into the treatment's potential impact on the male demographic and contribute significantly to our understanding of its effectiveness in this specific population.


## Females
```{r}
dd2 <- female_data %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

The box plot analysis, focusing on female patients receiving either the vaccine or placebo, reveals some intriguing findings. Among female patients who received the placebo, approximately 60% were infected with the covid91 virus, while those who received the vaccine had a lower infection rate of around 40%. Interestingly, both the placebo and vaccine groups showed a similar proportion of patients who remained uninfected, at approximately 50%. These observations raise questions about the factors influencing the infection rates in both groups and warrant further in-depth analysis to gain a clearer understanding of the underlying dynamics. Additional investigations and careful consideration of potential confounding variables are essential to elucidate the factors contributing to the differential infection rates between the placebo and vaccine groups in the female population.

### Numerical Descriptive Results
```{r}
table2 <- xtabs(~infected + treatment, data=female_data)
rowPerc(table2)
colPerc(table2)
```
The updated data presented in the contingency table pertains to female individuals and their infection status based on treatment (placebo and vaccine). Among female participants, 60% were infected with covid91 in the placebo group, while 40% were infected in the vaccine group. The overall distribution shows an equal split, with 50% of female individuals in each treatment group belonging to the OK category.

Analyzing the percentages within each treatment group, 2% of females in the placebo group were infected with covid91, leaving 98% uninfected. In contrast, in the vaccine group, 1.4% of females were infected with covid91, and 98.6% remained uninfected. These findings suggest a higher proportion of covid91 infections among females in the placebo group compared to those in the vaccine group.

However, to assess the significance of this discrepancy and establish any potential causal relationship between the treatment and infection status specifically in the female population, further rigorous statistical analysis would be required. The observed data provides valuable initial insights, but drawing definitive conclusions would necessitate applying appropriate statistical tests to evaluate the significance of the observed differences.

## Inferential Results
```{r}
chisq.test(table2)
chisqtestGC(table2)
femalefish <- fisher.test(table2)
femalefish
```
In the female patients, a comprehensive statistical analysis was performed to investigate the relationship between treatment and infection status. Pearson's Chi-squared test with Yates' continuity correction was conducted on the data, revealing a significant association between the treatment (placebo and vaccine) and infection status (covid19 and OK) in females. The test yielded a chi-square statistic of 13 with 1 degree of freedom and an impressively low p-value of 0.0003, indicating strong evidence against the null hypothesis. Additionally, Fisher's Exact Test for Count Data was applied, producing a similarly small p-value of 0.0002 and a 95% confidence interval for the odds ratio between 1.2 and 1.9.

### Effectiveness Confidence Interval 

```{r}
effect <- femalefish$p.value/(femalefish$p.value +1)
effectivenessfemale = femalefish$estimate/(femalefish$estimate +1)*100
print(effectivenessfemale)
```

```{r}
low=(femalefish$conf.int[1]/(femalefish$conf.int[1] +1))*100
hi=(femalefish$conf.int[2]/(femalefish$conf.int[2] +1))*100
options(digits=2)
```
#### Effectiveness Confidence Interval `r low`  to `r hi`


## LGBTQ



## Druggies


# Overall Results and Conclusions